

```{r}
library(data.table)
library(tidyverse)
library(parallel)
#load R functions
sapply(list.files("R/", ".r", full.names = TRUE), source)
set.seed(0203)
options(scipen = 10^6)


library(INLA)
library(fixest)
```

Prediction intervals are complicted with glm
We can get SE & confidence intervals for the expectation. 
Given mu, not obvious how to define the prediction interval
the prediction error should be  esier?


Models to be fitted
# warning
depite the name , inla predict Y (the count, directly), except , posterior sample are log count
BUT twfe is actually predicting smr
(becouse of the offset place)
```{r}
# this parametric bootstrap does rely on iid observations
# should implement a block bootstrap version
# stratified by year of adoption
# how woudl I handle non integers?

#summary(dat$Y0)
fit_twfe <- function(dat, nsim = 400) {
  t0 <- Sys.time()
  fm <- formula(Y0 ~  offset(log(E)) + x | fid + fyr)
  m0 <- fepois(fm
    , data = dat
    , fixef.rm = "none"
  )
  # predict : predict lambda
  pred <- predict(m0, newdata = dat)
  #summary(dat$Y)
  #summary(pred)
  #summary(exp(pred))
  #plot(log(dat$Y/dat$E), log(pred) )
  #round(sum(pred))
  #sum(dat$E)
  #sum(dat$Y)
#  summary(dat$E)

  # parametric bootstrap
  # simulate new residuals (Y0 - E * mu)
  # plot(resid(m0))
  dat[, paste0("SMR0.", 1:nsim)] <- sapply(1:nsim, \(i) {
    bdat <- dat
    bdat$Y0 <- rpois(length(pred), pred * bdat$E)
    #sum(bdat$Y0)
    #sum(bdat$Y)
    bdat$Y0[bdat$O != 1] <- NA
    mb <- suppressMessages(
      fepois(fm
        , data = bdat
        , fixef.rm = "none"
      )
    )
    pred <- predict(mb, newdata = bdat)
    pred
    #plot(log(bdat$Y / bdat$E), log(pred))
    #round(sum(bdat$Y / bdat$E))
    #round(sum(pred))
  })

  dat$SMR0.mean <-   pred
  dat$SMR0.sd   <-   apply(dat[, paste0("SMR0.", 1:nsim)], 1, sd)
  dat$method <- "twfe"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}


fit_inla_fe <- function(dat) {
  t0 <- Sys.time()
  m0 <- inla(Y0 ~ fid + fyr
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd
  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  #median(dat$SMR0.mean)
  #median(exp(dat$SMR0.1))
  dat$method <- "ife"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}


#this is equivalent to the syhtetic estiamtor
# in the sense that the fitted values from a "global" regression
# fitted to the sample, are used to estimate the level in eahc unit-year
# we do not take the average per-unit year,
# becouse there is a single observation
# the bias depends on the extent that the regression coeffcients are not homogenous
# note that we only explore heterogeneity across units.
# we could estimate the  bias taking the average residual per unit.
fit_inla_po <- function(dat) {
  t0 <- Sys.time()
  m0 <- inla(Y0 ~ x
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  #summary(dat$SMR0.mean)
  #summary(dat$Y)
  #plot(log(dat$Y / dat$E), log(dat$SMR0.mean / dat$E))
  #plot(log(dat$Y / dat$E), log(dat$SMR0.mean / dat$E))
  #plot(log(dat$Y / dat$E), log(exp(dat$SMR0.1) / dat$E))
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd

  sum(dat$SMR0.mean/dat$E)
  sum(exp(dat$SMR0.1))
  sum(dat$E)
  sum(dat$Y)

  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  dat$method <- "po"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}

fit_inla_ov <- function(dat) {
  dat$ii <- 1:nrow(dat)
  t0 <- Sys.time()
  m0 <- inla(Y0 ~ x #+ mxi + mxt
    + f(ii, model = "iid")
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd
  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  dat$method <- "ov"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}

fit_inla_ri <- function(dat) {
  t0 <- Sys.time()
  m0 <- inla(Y0 ~ x #+ mxi + mxt
    + f(id, model = "iid")
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd
  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  dat$method <- "ri"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}

fit_inla_re <- function(dat) {
  t0 <- Sys.time()
  m0 <- inla(Y0 ~ x #+ mxi + mxt
    + f(id, x, model = "iid")
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd
  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  dat$method <- "re"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}

# RE + AR(1)
fit_inla_ar <- function(dat) {
  t0 <- Sys.time()
  m0 <- inla(Y0 ~ x #+ mxi + mxt
    + f(id, model = "iid"
      , group = time, control.group = list(model = "ar1")
    )
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd
  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  dat$method <- "ar"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}

# spatial RE
fit_inla_sp <- function(dat, inla.graph = graph) {
  t0 <- Sys.time()
  m0 <- inla(Y0 ~ x #+ mxi + mxt
    + f(id, x, model = "besagproper2", graph = graph
      #, scale.model = TRUE
      #, group = time, control.group = list(model = "iid")
    )
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd
  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  dat$method <- "sp"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}

# full spatio-temporal model
fit_inla_st <- function(dat, inla.graph = graph) {
  t0 <- Sys.time()
  m0 <- inla(Y0 ~ x #+ mxi + mxt
    + f(id, x, model = "besagproper2", graph = graph
      #, scale.model = TRUE
      , group = time, control.group = list(model = "ar1")
    )
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd
  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  dat$method <- "st"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}

#separate components
fit_inla_fs <- function(dat, inla.graph = graph) {
  t0 <- Sys.time()
  dat$idx <- dat$id
  dat$ii <- 1:nrow(dat)
  m0 <- inla(Y0 ~ x
    #+ f(ii, model = "iid")
    + f(id, x, model = "besag", graph = graph
      , group = time, control.group = list(model = "iid")
    )
    , E = E
    , family = "poisson"
    , data = dat
    , control.predictor = list(link = 1)
    , control.compute = list(config = TRUE)
  )
  dat$SMR0.mean <-  m0$summary.fitted.values$mean
  dat$SMR0.sd   <-  m0$summary.fitted.values$sd
  ps <- inla.posterior.sample(400, m0)
  dat[, paste0("SMR0.", 1:400)] <- inla.posterior.sample.eval("Predictor", ps)
  dat$method <- "fst"
  t1 <- Sys.time()
  print(t1 - t0)
  dat
}
```

DATA
```{r}
dt_list <- readRDS(file = paste0("data/placebo.rds"))
set.seed(0203)
dat <- dt_list[[5]]
dat$Y <- dat$Y0 <- round(dat$SMR * dat$E, 0)
dat$Y0[dat$O != 1] <- NA
dat$id   <- as.numeric(factor(dat$fid))
dat$time <- as.numeric(as.factor(dat$fyr))
head(dat)


dat <- dat %>%
  group_by(id) %>%
  mutate(mxi = mean(x)) %>%
  group_by(time) %>%
  mutate(mxt = mean(x)) %>%
  ungroup
```

nonspatial
```{r}
twfe <- fit_twfe(dat)

sum(twfe$SMR0.1)
sum(twfe$SMR0.mean)
#mvc <- fit_vc(dat)
#mfe <- fit_inla_fe(dat)
mpo <- fit_inla_po(dat)
mov <- fit_inla_ov(dat)
mri <- fit_inla_ri(dat)
mre <- fit_inla_re(dat)
mar <- fit_inla_ar(dat)
```
spatial
```{r}
load("data/gs")
graph <- gs
set.seed(0203)

msp <- fit_inla_sp(dat)
mst <- fit_inla_st(dat)

```

#Dawid–Sebastiani score
#depedns only on mean and SE (or poserior SD)
```{r}
dss <- \(m) {
  m %>%
    filter(O == 0) %>%
    mutate(
      lambda = SMR0.mean * E
      , v_lamb = (SMR0.sd * E) ^ 2
      , sigma = sqrt(lambda + v_lamb)
    ) %>%
    reframe(dss = mean(
      ((Y - lambda) / sigma) ^ 2 + 2 * log(sigma)
    ))
}

  twfe %>%
    filter(O == 0) %>%
    mutate(
      lambda = SMR0.mean * E
      , v_lamb = (SMR0.sd * E) ^ 2
      , sigma = sqrt(lambda + v_lamb)
    ) %>%
    mutate(dss = 
      ((Y - lambda) / sigma) ^ 2 + 2 * log(sigma)
    ) %>% select(fid, fyr, SMR0.mean, SMR0.sd, sigma, dss, Y) %>% .[45:56, ]

# imagine a case with no suicides  before the policy
# the fe is zero, and the varaince is close to zero (a lot of certainty)
# but there is one case after the policy 
# then the ddiscrepancy between observed and expected, RELATIVE to the se,  is  hughe 


bind_rows(mvc, twfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  dss()


bind_rows(twfe, mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method, urb) %>%
  dss() %>%
  select(!o) %>%
  pivot_wider(names_from = urb, values_from = dss)
```

how good are the point  estimates
```{r}
corp <- \(m) {
  m %>%
    filter(O == 0) %>%
    reframe(cor = cor(Y, SMR0.mean * E))
}


bind_rows(twfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  corp()
bind_rows(twfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method, urb) %>%
  corp() %>%
  select(!o) %>%
  pivot_wider(names_from = urb, values_from = cor)

cork <- \(m) {
  m %>%
    filter(O == 0) %>%
    reframe(cork = cor(rank(Y), rank(SMR0.mean * E)))
}

bind_rows(twfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  cork()
bind_rows(twfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method, urb) %>%
  cork() %>%
  select(!o) %>%
  pivot_wider(names_from = urb, values_from = cork)

rss <- \(m) {
  m %>%
    filter(O == 0) %>%
    reframe(
      rss = mean((Y - SMR0.mean * E)^2 / (SMR0.mean * E))
      )
}


bind_rows(twfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  rss()


```


```{r}
sse_smr <- \(m) {
  m %>%
    filter(O == 0) %>%
    reframe(sse_smr = mean(
      mean((SMR - SMR0.mean) ^ 2)
    ))
}
ses <- \(m) {
  m %>%
    filter(O == 0) %>%
    mutate(
      lambda = SMR0.mean * E
    ) %>%
    reframe(ses = mean(
      mean((Y - lambda)^2)
    ))
}
aes <- \(m) {
  m %>%
    filter(O == 0) %>%
    mutate(
      lambda = SMR0.mean * E
    ) %>%
    reframe(aes = mean(
      abs((Y - lambda))
    ))
}

bind_rows(twfe, mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  sse_smr()

bind_rows(mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method, urb) %>%
  sse_smr() %>%
  select(!o) %>%
  pivot_wider(names_from = urb, values_from = sse_smr)


bind_rows(twfe, mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  ses()

bind_rows(twfe,mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method, urb) %>%
  ses() %>%
  select(!o) %>%
  pivot_wider(names_from = urb, values_from = ses)


bind_rows(twfe, mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  aes()

bind_rows(twfe, mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method, urb) %>%
  aes() %>%
  select(!o) %>%
  pivot_wider(names_from = urb, values_from = aes)

```


*** Becouse the true att is zero att = bias
```{r}
att <- \(m) {
  m %>%
    filter(O == 0) %>%
    mutate(
      lambda = SMR0.mean * E
    ) %>%
    reframe(att = mean(
      (Y - lambda)
    ))
}
att(twfe)
att(mov)*1
att(mfe)*1
att(mpo)

bind_rows(twfe, mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  att()

bind_rows(mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method, urb) %>%
  att() %>%
  select(!o) %>%
  pivot_wider(names_from = urb, values_from = att)

colnames(mre)

att_sd <- \(m) {
  m %>%
    filter(O == 0) %>%
    mutate(
      across(SMR0.1:SMR0.200, \(x) x * E)
    ) %>%
    reframe(
      across(SMR0.1:SMR0.200, mean)
    ) %>%
    rowwise() %>%
    reframe(
      att_sd = sd(c_across(SMR0.1:SMR0.200))
    )
}

att(mfe)
att_sd(mfe)

bind_rows(twfe, mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  att

bind_rows(twfe, mfe, mpo, mri, mre, mar, msp, mst, .id = "o") %>%
  group_by(o, method) %>%
  att_sd


```


```{r}
load("data/gs")
graph <- gs
set.seed(0203)

fit_all <- \(dat) {
  twfe <- fit_twfe(dat)
  mpo <- fit_inla_po(dat)
  mri <- fit_inla_ri(dat)
  mre <- fit_inla_re(dat)
  mar <- fit_inla_ar(dat)
  #spatial
  msp <- fit_inla_sp(dat)
  mst <- fit_inla_st(dat)
  bind_rows(twfe, mpo, mri, mre, mar, msp, mst, .id = "o")
}

start_time <- Sys.time()
mclapply(names(dt_list), \(name) {
  print(name)
  dt <- dt_list[name][[1]]
  dt$Y <- dt$Y0 <- round(dt$SMR * dt$E, 0)
  dt$Y0[dt$O != 1] <- NA
  dt$id   <- as.numeric(factor(dt$fid))
  dt$time <- as.numeric(as.factor(dt$fyr))
  rep <-  tryCatch(
    #fit_all(dt)
    fit_twfe(dt)
    ,  error = function(c) NULL
  )
  saveRDS(rep, file = paste0("data/tmp/placebo_rep_", name))
}, mc.cores = 10
)
end_time <- Sys.time()
end_time - start_time

list_files <- list.files("data/tmp", "placebo_rep_", full.names = TRUE)
list_dat   <- lapply(list_files, \(fl) readRDS(fl))
length(list_dat)

#fits <- fits %>%
# mutate(SMR0.sd = if_else(is.na(SMR0.sd),SMR0.sd2,SMR0.sd)
## )

fits      <- rbindlist(list_dat, idcol = "replication")
#saveRDS(fits, file = paste0("data/placebo_rep.rds"))
saveRDS(fits, file = paste0("data/placebo_twfe.rds"))

```

```{r}
all_m <- \(m) {
  m %>%
    filter(O == 0) %>%
    mutate(
      lambda = SMR0.mean * E
      , v_lamb = (SMR0.sd * E) ^ 2
      , sigma = sqrt(lambda + v_lamb)
    ) %>%
    reframe(
      dss = mean(((Y - lambda) / sigma) ^ 2 + 2 * log(sigma))
      , att = mean((Y - lambda))
      , ses = mean(mean((Y - lambda)^2))
      , aes = mean(abs((Y - lambda)))
      , sse_smr = mean(mean((SMR - SMR0.mean) ^ 2))
      , across(Y0.1:Y0.400, mean)
      , Y = mean(Y)
    )
}

```


```{r}
fits <- readRDS(file = paste0("data/placebo_rep.rds"))
dim(fits)
fits <- filter(fits, O == 0)
dim(fits)
set.seed(0203)
dim(fits)

sim_count <- \(x) rpois(length(x), exp(x) * fits$E)
fits[, (paste0("Y0.", 1:400)) := lapply(.SD, sim_count)
     , .SDcols = paste0("SMR0.", 1:400)]

#fits[, (paste0("att.", 1:400)) := lapply(.SD, \(x) fits$Y - x)
#     , .SDcols = paste0("Y0.", 1:400)]

colnames(fits)

#fits[,  att_sd := apply(.SD, 1, sd)
#     , .SDcols = paste0("Y0.", 1:400)]

#fits[,  att_lw := apply(.SD, 1, quantile, .025, na.rm = TRUE)
#     , .SDcols = paste0("Y0.", 1:400)]

#fits[,  att_up := apply(.SD, 1, quantile, .975, na.rm = TRUE)
#     , .SDcols = paste0("Y0.", 1:400)]

#fits[,  cov :=  as.numeric(att_lw < 0 & att_up > 0)]
```

```{r}
#fits$replication[1:10]
#test <- filter(fits, replication == "1" & method == "re" & O == 0)
#dim(test)
#head(test)
#summary(test$)


pl_tb <- fits %>%
  group_by(o, method, replication) %>%
  all_m()

filter(pl_tb, method == "re") %>%
  mutate(across(Y0.1:Y0.200, \(x) Y - x)) %>%
  rowwise() %>%
  mutate(
    sd = sd(c_across(Y0.1:Y0.200))
    , lw = quantile(c_across(Y0.1:Y0.200), .025, na.rm = TRUE)
    , up = quantile(c_across(Y0.1:Y0.200), .975, na.rm = TRUE)
  ) %>%
  select(method, replication, att, lw, up, sd, Y, Y0.1:Y0.20)


  %>%
  rowwise() %>%
  mutate(
    att_sd = sd(c_across(Y0.1:Y0.400))
    , att_up = Y - quantile(c_across(Y0.1:Y0.400), .025, na.rm = TRUE)
    , att_lw = Y - quantile(c_across(Y0.1:Y0.400), .975, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  mutate(
    cov = as.numeric(att_lw < 0 & att_up > 0)
  ) %>%
  group_by(o, method) %>%
  reframe(
    across(c(dss:aes, att_sd, cov), list(mean, sd))
  )

select(pl_tb, method, att_1, att_2, cov_1)



```